---
title: "Collapsing Cabi data"
author: "Bingo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

```{r read}
library(arrow)
library(tidyverse)
# load this thing in

cabi <- arrow::read_parquet("C:/Users/bingo/OneDrive - Georgia Institute of Technology/CSE7643/project/cabi_combined_data.parquet")


```

# Data engineering

First we need to split the date columns into date and the time. We want to know what time these things happened at. 

```{r convert_date}
cabi <- cabi %>% 
  mutate(started_at = lubridate::as_datetime(started_at),
         ended_at = lubridate::as_datetime(ended_at))
```

Now that they are dates, lets bucket them into 15 minute intervals

```{r}
cabi <- cabi %>% 
  mutate(started_at_15 = lubridate::floor_date(started_at, unit = "15 minutes"),
         ended_at_15 = lubridate::floor_date(ended_at, unit = "15 minutes"))
```

Now we can group by the station and the 15 minute intervals and make 2 datasets. One for leavings and one for enterings.

```{r}
departures <- cabi %>% 
  group_by(start_station_id, started_at_15) %>% 
  summarise(deps = n())
arrivals <- cabi %>% 
  group_by(end_station_id, ended_at_15) %>% 
  summarise(arrs = n())
```

Now lets join them together and calculate demand.

```{r}
deps_and_arrs <- departures %>% 
  full_join(arrivals, by = join_by(x$start_station_id == y$end_station_id, x$started_at_15 == y$ended_at_15)) %>% 
  mutate(demand = arrs - deps)
```



# Analysis

Our goal here is to compute the following:
compute net arrivals - departures by 15 minute intervals



# write out

```{r}
deps_and_arrs %>% 
  arrow::write_parquet(., "C:/Users/bingo/OneDrive - Georgia Institute of Technology/CSE7643/project/cabi_demand_15.parquet")
```


